DOM Review Summary

### 2024-03-18 Analysis/ReviewSummary/ReviewSummary.qmd
study size over time
plotted trends/boxplots for study size over time

Study duration, defined as the length of time between the first and last survey, ranged from **`{r} min(dataCollection_clean$Duration)`** month to **`{r} max(dataCollection_clean$Duration)/12`** years, with a median of **`{r} median(dataCollection_clean$Duration`** months.

@@

dataCollection_clean |>
  mutate(Total = length(unique(ID_clean))) |>
  separate_rows(Detection_Method, sep = ", ") |>
  summarise(Articles = n(), 
            #Proportion = round(Articles/Total, 2),
            .by = c(Detection_Method, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  select(Detection_Method, Articles, Percentage) |>
  filter(!Detection_Method %in% c("66", NA, "Unspecificed",
                                  "Survey", "Incidental")) |>
  arrange(Percentage) |>
  mutate(Detection_Method = case_when(Detection_Method == "AudioVisual" ~ "Human structured survey",
                                      Detection_Method == "Search" ~ "Human unstructured search",
                                      Detection_Method == "Trapping" ~ "Physical trapping",
                                      Detection_Method == "Camera" ~ "Camera trapping",
                                      Detection_Method == "Citizen" ~ "Citizen science",
                                      Detection_Method == "Bioacoustic" ~ "Acoustic monitor")) |>
  mutate(Detection_Method = fct_inorder(Detection_Method)) |>

@@

```{r DataCollection_Sampling}

@@

>>>>

Study duration, defined as the length of time between the first and last survey, ranged from **`{r} min(dataCollection_clean$Duration)`** month to **`{r} max(dataCollection_clean$Duration)/12`** years, with a median of **`{r} median(dataCollection_clean$Duration)`** months.

@@

dataCollection_clean |>
  mutate(Total = length(unique(ID_clean))) |>
  separate_rows(Detection_Method, sep = ", ") |>
  summarise(Articles = n(), 
            .by = c(Detection_Method, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  select(Detection_Method, Articles, Percentage) |>
  filter(!Detection_Method %in% c("66", NA, "Unspecificed",
                                  "Survey", "Incidental")) |>
  arrange(Percentage) |>
  mutate(Detection_Method = case_when(Detection_Method == "AudioVisual" ~ "Human structured survey",
                                      Detection_Method == "Search" ~ "Human unstructured search",
                                      Detection_Method == "Trapping" ~ "Physical trapping",
                                      Detection_Method == "Camera" ~ "Camera trapping",
                                      Detection_Method == "Citizen" ~ "Citizen science",
                                      Detection_Method == "Bioacoustic" ~ "Acoustic monitor")) |>
  mutate(Detection_Method = fct_inorder(Detection_Method)) |>

@@

Site selection was variable; various studies selected sites randomly, as part of a grid design, preferentially, or exhaustively. A considerable amount did not clearly specify how their study sites were determined.

```{r DataCollection_Sampling}
#| echo: false

dataCollection_clean |>
  mutate(Total = length(unique(ID_clean))) |>
  separate_rows(Sampling_Method, sep = ", ") |>
  mutate(Sampling_Method = case_when(Sampling_Method == "Unspecified" ~ "Not specified",
         .default = Sampling_Method)) |>
  summarise(Articles = n(), 
            .by = c(Sampling_Method, Total)) |>
  mutate(Percentage = round(Articles/Total, 2)*100) |>
  select(-Total) |>
  arrange(-Percentage) |>
  kable()
```

@@

There does not appear to be a discernible trend in study size (whether defined by duration, number of sites, or total number of survey occasions.

::: panel-tabset
## Year

```{r DataCollection_sizeTrend}
#| echo: false
#| warning: false
#| message: false

dataCollection_clean |>
  select(ID_clean, Duration, 
         Primary_Max, Secondary_Max,
         SiteNumber_Max) |>
  mutate(ID_clean = as.numeric(ID_clean)) |>
  rowwise() |>
  mutate(TotalOccasions = Primary_Max*Secondary_Max) |>
  left_join(select(studyDetails_clean, ID, Year), by = join_by(ID_clean == ID)) |>
  pivot_longer(cols = c(Duration, TotalOccasions, SiteNumber_Max),
               names_to = "Variable", values_to = "Quantity") |>
  mutate(Variable = case_when(Variable == "Duration" ~ "Study duration (months)",
                              Variable == "SiteNumber_Max" ~ "Number of sites",
                              Variable == "TotalOccasions" ~ "Total survey occasions (Primary x Secondary)")) |>

  ggplot(aes(x = Year, y = Quantity, colour = Variable)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, alpha = 0.7) +
  scale_y_continuous(transform = "log10") +
  xlim(c(2005, 2024)) +
  scale_x_continuous(breaks = c(2005, 2010, 2015, 2020, 2024)) +
  scale_colour_manual(values = c("darkslategrey", "cyan4", "darkorchid4")) +
  facet_wrap(vars(Variable), ncol = 1, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Trend in study size",
       x = "Year", y = "")
```

## Year Strata

```{r DataCollection_Sizetrendboxplots}
#| echo: false
#| warning: false
#| message: false

dataCollection_clean |>
  select(ID_clean, Duration, 
         Primary_Max, Secondary_Max,
         SiteNumber_Max) |>
  mutate(ID_clean = as.numeric(ID_clean)) |>
  rowwise() |>
  mutate(TotalOccasions = Primary_Max*Secondary_Max) |>
  left_join(select(studyDetails_clean, ID, YearStrata), by = join_by(ID_clean == ID)) |>
  filter(!is.na(YearStrata)) |>
  pivot_longer(cols = c(Duration, TotalOccasions, SiteNumber_Max),
               names_to = "Variable", values_to = "Quantity") |>
  mutate(Variable = case_when(Variable == "Duration" ~ "Study duration (months)",
                              Variable == "SiteNumber_Max" ~ "Number of sites",
                              Variable == "TotalOccasions" ~ "Total survey occasions (Primary x Secondary)")) |>

  ggplot(aes(x = YearStrata, y = Quantity, colour = Variable)) +
  geom_point(alpha = 0.3) +
  geom_boxplot() +
  scale_y_continuous(transform = "log10") +
  scale_colour_manual(values = c("darkslategrey", "cyan4", "darkorchid4")) +
  facet_wrap(vars(Variable), ncol = 1, scales = "free_y") +
  theme(legend.position = "none") +
  labs(title = "Trend in study size",
       x = "Year strata", y = "")
```
:::
